\addcontentsline{toc}{chapter}{Bibliografía}
\begin{thebibliography}{99}

\bibitem{antigua}
\textit{C. Ruggles and G. Urton, *Skywatching in the Ancient World: New Perspectives in Cultural Astronomy*. Boulder, CO: University Press of Colorado, 2007.}

\bibitem{maya} \textit{Stanislaw Iwaniszewski, Arqueología Mexicana, "Los ciclos lunares y el calendario maya".}  [Online]. Disponible: \url{https://arqueologiamexicana.mx/mexico-antiguo/los-ciclos-lunares-y-el-calendario-maya}

\bibitem{ixchel} \textit{C. Carcavilla, "Trabajo de Fin de Grado Arqueología UCM: La diosa Ix Chel en los mayas prehispánicos", 2022} Disponible: \url{https://www.researchgate.net/publication/363582211_Trabajo_de_Fin_de_Grado_Arqueologia_UCM_La_diosa_Ix_Chel_en_los_mayas_prehispanicos}.

\bibitem{sidereus} \textit{G. Galilei, *Sidereus Nuncius*. Venice, Italy: Thomas Baglioni, 1610. [Online]. Available}: \url{https://archive.org/details/sidereus_nuncius_202004/page/n9/mode/2up}. [Accessed: Feb. 9, 2025].

\bibitem{bbc} \textit{BBC Mundo, "La sorprendente revelación sobre la Luna y su evolución", Dec. 2, 2017. [Online]. Available:} \url{https://www.bbc.com/mundo/noticias-42202220}. [Accessed: Feb. 9, 2025].

\bibitem{decaelo} \textit{A. Magno and H. Mitanuni, *De caelo et mundo*. Venice, Italy: Impressum Uenetijs per Joannem de Forliuio, et Gregorium fratres, Nov. 1490. [Online]. Available:} \url{http://hdl.handle.net/10481/16176}. [Accessed: Feb. 9, 2025].

\bibitem{lunarspine} ResearchGate, 
``The influence of lunar phases and zodiac sign ‘Leo’ on perioperative complications and outcome in elective spine surgery,'' 
Scientific Figure, 2025. Disponible en: \href{https://www.researchgate.net/figure/Woodcut-from-1492-illustrating-the-lunar-phases-from-the-chapter-Liber-de-coelo-et_fig1_301581818}{ResearchGate}. [Accedido: 10 Feb. 2025].


\bibitem{huijse2014} P. Huijse, P. A. Estevez, P. Protopapas, J. C. Principe, y P. Zegers, 
``Computational Intelligence Challenges and Applications on Large-Scale Astronomical Time Series Databases,'' 
\textit{IEEE Computational Intelligence Magazine}, vol. 9, no. 3, pp. 27–39, 2014. \url{https://arxiv.org/pdf/1509.07823}

\bibitem{jarrett2021} T. H. Jarrett, A. Comrie, L. Marchetti, A. Sivitilli, S. Macfarlane, F. Vitello, U. Becciani, A. R. Taylor, J. M. van der Hulst, P. Serra, N. Katz y M. E. Cluver, 
``Exploring and interrogating astrophysical data in virtual reality,'' 
\textit{Astronomy and Computing}, vol. 37, p. 100502, 2021. 
DOI: \url{https://doi.org/10.1016/j.ascom.2021.100502}.

\bibitem{aprendizajeEscuelas} \textit{Solbes and R. Palomar, “Dificultades en el aprendizaje de la astronomía en secundaria”, Rev. Bras. Ensino Fís., vol. 35, no. 1, pp. 01–12, Jan. 2013, doi: 10.1590/S1806-11172013000100016.}

\bibitem{rabanales2021} F. Rabanales Loyola y C. Vanegas Ortega, 
``Concepciones alternativas sobre astronomía en estudiantes de educación básica y media de la Región Metropolitana de Chile,'' 
\textit{Estudios Pedagógicos (Valdivia)}, vol. 47, no. 2, pp. 247–268, 2021.

\bibitem{conceptos2009} J. Alfonso Garzón, D. Galadí Enríquez, y C. Morales Durán, Eds., \textit{100 Conceptos Básicos de Astronomía. Madrid, España: Instituto Nacional de Técnica Aeroespacial (INTA), 2009}

\bibitem{iau2025} Marissa Rosenberg, Pedro Russo, Unión Astronómica Internacional, 
\textit{Astronomía en la vida cotidiana}. [En línea]. Disponible en: \href{https://www.iau.org/public/themes/astronomy_in_everyday_life/spanish/}{https://www.iau.org/public/themes/astronomy\_in\_everyday\_life/spanish/}.

\bibitem{gullari2025} J. K. Gullari, 
\textit{The Impact of Moon Phases on Earth, Plants, and Humans: A Comprehensive Study from Project Alpha}, 
Jan. 9, 2025.

\bibitem{Singh2022}
A.~Singh, A.~Sharma, S.~Kumar, L.~Chang, A.~Vashishtha, R.~Raj, S.~Agrawal, and P.~Chauhan, ``\textit{Dielectric characterization and polarimetric analysis of lunar north polar crater Hermite-A using Chandrayaan-1 Mini-SAR, Lunar Reconnaissance Orbiter (LRO) Mini-RF, and Chandrayaan-2 DFSAR data},'' \textit{Advances in Space Research}, vol.~70, no.~12, pp.~4030--4055, 2022. doi: 10.1016/j.asr.2022.04.059.


\bibitem{ondas} \textit{L. D. Herrera and C. R. Arias Simulación de ondas generadas en la inestabilidad two-stream para
plasmas electrostáticos usando particle-in-cell simulación de ondas generadas en la inestabilidad
two-stream para plasmas electrostáticos usando particle-in-cell, 2019.}

\bibitem{perryman}\textit{M. Perryman, The Exoplanet Handbook. 2011. [Online]. Available: } \url{https://assets.cambridge.org/97805217/65596/frontmatter/9780521765596_frontmatter.pdf}


\bibitem{green} \textit{Greenfield, P., “What Python Can Do for Astronomy”, in <i>Astronomical Data Analysis Software and Systems XX</i>, 2011, vol. 442, p. 425.}

\bibitem{PerezLisboa2020}
S.~Pérez-Lisboa, C.~G. Ríos-Binimelis, and J.~Castillo Allaria, ``\textit{Realidad Aumentada y simuladores: astronomía para niños y niñas de cinco años},'' \textit{ALTERIDAD. Revista de Educación}, vol.~15, no.~1, pp.~25--35, 2020. doi: 10.17163/alt.v15n1.2020.02.


\bibitem{nasa} \textit{Leonardi, L., Daricello, L., and Giacomini, L., “Learning astronomy through Augmented Reality: EduINAF resources to enhance students' motivation and understanding”}  Art. no. EPSC2021-530, 2021. doi:10.5194/epsc2021-530.

\bibitem{Ucar2014}
S.~Uçar, ``\textit{The effects of simulation-based and model-based education on the transfer of teaching with regard to Moon phases},'' \textit{Journal of Baltic Science Education}, vol.~13, no.~3, pp.~327--338, 2014.

\bibitem{UNL2025} 
University of Nebraska-Lincoln, 
``\textit{Lunar Phases Simulator},'' 
[Online]. Available: \href{https://astro.unl.edu/naap/lps/animations/lps.html}{https://astro.unl.edu/naap/lps/animations/lps.html}. 
[Accessed: 11-Feb-2025].


\end{thebibliography}